---
title: "Homework-4"
author: "Mary Cadogan"
format: pdf
editor: visual
---

1.  Mathematical: H0: β~1~ = 0

    HA: β~1~ ≠ 0

    Biological: H0: Fish length does not predict fish weight for trout perch.

    HA: Fish length does predict fish weight for trout perch.

```{r}
#load in packages
library(tidyverse)
library(here)
library(naniar)
library(flextable)
library(performance)
library(broom)
library(car)
library(ggeffects)
```

```{r}
#read in the data and create new object
fish_data <- read_csv(here("code", "ntl6_v12.csv"))
```

2.  

```{r}
#filtered data set by troutperch, and then selected for the 2 variables we are looking at: length and weight
troutperch_summary <- fish_data %>% 
  filter(spname == "TROUTPERCH") %>% 
  select("length", "weight") 
```

```{r}
#see if any weight or length data is missing using gg_miss_var
gg_miss_var(troutperch_summary) %>% 
  labs(caption = "The missing data from the weight column will impact the hypotheses due to weight being the dependent variable we are testing. 200 missing variables is a substantial amount of data that cannot be utilized to determine a linear relationship. There was no missing data for fish lengths. ")
  
```

```{r}
#get a general sense of what the data looks like using a point graph
ggplot(data = troutperch_summary, aes(x = length, y = weight)) + 
  geom_point() 
```

3\.

```{r}
#run linear model test with length affecting weight from filtered data set 
troutperch_object <- lm(weight ~ length, data = troutperch_summary)
summary(troutperch_object)
```

4\.

```{r}
#show results of linear modeling tests visually in 4 different plots in a grid
par(mfrow =c(2,2))
plot(troutperch_object)

#check other assumptions
check_heteroscedasticity(troutperch_object)

check_normality(troutperch_object)
```

5.  Residuals vs Fitted: Most of the data for the linear modeling test is distributed evenly around a relatively horizontal line but with 3 identified outliers. Therefore, normality is maintained but not homoscedasticy.

Normal Q-Q: The data points are in a fairly straight line, except for the 3 outliers at the end. Therefore, the data could be considered not normally distributed and potentially heteroscedastic due to the outliers.

Scale-Location: The Scale-Location is similar to the results from Residuals vs Fitted, with a similar horizontal line besides the outliers. Therefore normality is maintained but not homoscedasticy.

Residuals vs Leverage: This plot uses Cook's distance to determine if any outliers have a significant influence on the data. This plot shows that 226 might have a large influence on the data.

I decided to keep all data points because I did not think they would heavily impact the slope of the line and the data points may be plausible instead of a "mistake".

6\.

```{r}
trout_summarytable <- summary(troutperch_object)

trout_summarytable
```

7\.

```{r}
trout_anova <-anova(troutperch_object)

trout_anova_table <- tidy(trout_anova) %>% 
  mutate(across(sumsq:meansq, ~ round(.x, digits = 4))) %>%
  mutate(statistic = round(statistic, digits = 1)) %>% 
  mutate(p.value = case_when(p.value < 0.001 ~ "< 0.001")) %>%
  flextable() %>% 
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

trout_anova_table
```

8.  The ANOVA table relates to the information from the summary() object by showing the same (or very similar) F statistic, the p value, and the degrees of freedom.
9.  The results indicate that the data rejects the null hypothesis, that troutperch lengths do not predict weight, in favor of the alternative hypothesis, that troutperch lengths do predict weight. Testing a total of 489

```{r}
#use ggpredict to make a predictions plot
predictions <- ggpredict(troutperch_object, terms = "length")
predictions

#plotting predictions from the ggpredict, with length on x and weight on y
predict_plot <- ggplot(data = troutperch_object, 
                       aes(x = length, y = weight)) +
  #plotting the troutperch data lone
  geom_point() +
  #plotting predictions of linear model, changing aesthetics of plot line
  geom_line(data = predictions, 
              aes(x = x, y = predicted), 
              color = "blue", linewidth = 0.75) +
  #plotting the confidence interval from the gg predict function (95%)
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +

  #theme and meaningful labels, expand on caption
  theme_bw() +
  labs(x = "Length",
       y = "Weight", title = "Does Troutperch Length Predict Weight?", caption = "Linear model predictions determined using lm() test. Data source: North Temperate Lakes LTER: Fish Lengths and Weights (1981)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0.5))

predict_plot
```
